---
title: "Lab 5"
author: "Nicholas Stout"
date: "October 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
accident = read.csv("accident.csv")
person = read.csv("person.csv")
vehicle = read.csv("vehicle.csv")
library(tidyverse)
library(ggmap)
library(maps)
library(mapdata)
```

###Load the accident data, plot latitude and longitude on a map

```{r}
usa <- map_data("usa")
ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "red")+ geom_point(data=accident, aes(x=LONGITUD, y=LATITUDE)) + xlim(-175,-30)+ylim(19,70)
```

##Part 2
###Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.
###Join accident and driver table (work out which variable(s) to use)
###Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
drivers = person %>% filter(PER_TYP == 1)
drivers %>% select(-c(HOUR)) %>% left_join(accident, by="ST_CASE") %>% group_by(DAY_WEEK) %>% summarise(count = n()) %>% ggplot(aes(x=DAY_WEEK)) + geom_bar()

```

Days 1, 6, and 7 are higher than the other days of the week.
